\LARGE\singlespacing{\textbf{Appendix} \\ \medskip Air cleaners and respiratory infections in schools: \\ A modeling study using epidemiological, environmental, and molecular data}

### 2023-11-07 doc/paper/supplements.tex
Updates from Overleaf

\title{\LARGE\singlespacing{\textbf{S1 Appendix} \\ \medskip
Air cleaners and respiratory infections in schools: \\ A modeling study using epidemiological, environmental, and molecular data}}

@@

The clean air delivery rate (CADR) is typically used to quantify the cleaning efficiency of air purifiers\cite{Nelson1993IndoorEnv}. It is expressed in m$^3$/s and describes the volumetric flow rate of particle-free, clean air that an air purifier delivers into the room. The CADR is equal to the product of the filtration efficiency and the volumetric flow rate passing through the device.

We followed the approach by K\"{u}pper et al.\cite{Kuepper20199AAQR} to determine the CADR of the air purifiers used in this study. The approach is based on the release of fine aerosols (particle diameter $D$<1$\mu$m) in a closed, unventilated room by measuring the decay rate $k$ of the particle concentration $N(t)$ caused by the air cleaner, \ie
\begin{align}
    N(t) = N_0 \cdot e^{-k\cdot t}
\end{align}
Since the surfaces in the room (\eg walls and furniture) are also sinks for aerosol particles, measurements were carried out with ($k_\mathrm{cleaner}$) and without air cleaners($k_\mathrm{background}$). Based on that, we can compute the CADR as
\begin{align}
    \mathrm{CADR} = V \cdot k_\mathrm{cleaner} - V \cdot k_\mathrm{background},
\end{align}
where $V = 233 m^3$ is the volume of the classroom.

@@

Figure~\ref{fig:experiment} shows the particle number concentration with and without without the air purifier. The exponential decrease is clearly visible as a linear curve in the logarithmic plot. Based on these experiments, we were able to determine the net CADR values of the air purifier as 420\,m$^3$/h with an uncertainty of $\pm$10\% due to systematic differences in the setup and measurement errors of the instruments. Our estimated CADR is lower than the CADR reported by the manufacturer of 600m$^3$/h, most likely because the reported CADR was determined with comparably large particles (pollen in the super micrometer size range). Thus, we explain the lower filtration efficiency determined here with the fact that CADR is strongly dependent on the particle size, \ie large particles are easier to filter than fine aerosols.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{doc/paper/experiment-particle-concentration-decay.png}
    \caption{Measured concentration curves with the air purifier on and off. The red dot shows the average concentration measured by the 11 distributed sensors of particles with $0.3\mu m < D < 1 \mu m$. The standard deviation is shown in gray as an error bar. The increase in concentration at the beginning of these experiments is due to the release of the NaCl particles during approximately 15-20\,min. The inhomogeneity of the concentration during the nebulization period is seen in larger error bars. Afterwards, the aerosol source is turned off, the concentration homogenizes quickly, and the concentration drop follows an exponential decay, which is used for the CADR calculation.}
    \label{fig:experiment}
\end{figure}

@@

\begin{table}[!htpb]
    \footnotesize
    \centering
    \caption{\textbf{Data collection}. Type of data collected, method/device, and frequency in the rooms of classes A and B.}
    \begin{tabular}{p{3.5cm}p{6cm} p{1cm} p{1cm} p{3cm}}
    \midrule
    Type of data & Method / device & Class A & Class B & Frequency \\
    \midrule
    \textbf{Molecular data} \\
    \midrule
    Bioaerosol sampling & Bioaerosol sampling devices: BioSpot-VIVAS (Aerosol Devices Inc., Ft. Collins, Colorado, United States) and Coriolis Micro Air (Bertin Instruments Montigny-le/Bretonneux, France) & x & (x) \newline only Coriolis & daily \\
    Saliva samples & Saliva samples & x & x & twice per week \\
    Swabs from air cleaners & Swab samples from Xiaomi HEPA filters (Xiaomi Mi Air Pro 70m2, Shenzhen, China) & x & x & after each intervention phase (both classes) and before the vacation (only class B) \\ 
    \midrule
    \textbf{Environmental data} \\
    \midrule
    CO$_2$, aerosol/particle concentrations, humidity, temperature & Air quality device: AQ Guard (Palas GmbH, Karlsruhe, Germany) & x & x & daily by minute \\
    Coughs & Sounds collected via microphones and coughs detected with a Machine Learning model \cite{Bertschinger2023CBMS} & x & x & daily by seconds \\
    \midrule
    \textbf{Epidemiological data} \\
    \midrule
    Absences & Survey & x & x & daily \\
    Student characteristics & Survey & x & x & at study start \\
    \bottomrule
    \end{tabular}
    \label{tab:data}
\end{table}

@@

Reports about absences were entered electronically into a REDcap database\cite{Harris2009,Harris2019}. \Cref{tab:epi-data-line-list} shows the line list data for respiratory cases based on epidemiological data. For each class and case, it shows the date when the student was absent, s/he reported symptoms, and returned to school. We also collected data on laboratory tests but none of the students reported a test result during the study. A case of respiratory infection was defined as an absence where the student reported a sickness with at least one of the following symptoms: fever, coughing, tiredness, loss of test or smell, sore throat, headache, aches and pains, diarrhea, difficulty breathing or shortness of breath, stomach. Students were asked to report the first date they experienced symptoms, so that a student absent on Monday could report Saturday or Sunday as the day when symptoms began to be felt. We will refer respiratory cases by their dates of symptom onset, which usually corresponded to the absence date, unless the student attended the school while experiencing symptoms. Informed consent for one student could not be obtained and the single absence of this student was considered unrelated to a respiratory infection.

@@

The prior distributions for the virus-specific incubation periods are based on estimates published in the literature\cite{McAloon2020,Lessler2009LancetID}. The distributions are shown in \Cref{fig:prior-pin}. Since we could not obtain prior estimates for the incubation period of metapneumovirus (MPV) from the literature, we instead formed a distribution by using the equally weighted average of the parameters from the other distributions.

@@

We use weakly informative priors for all modeling parameters following recommendations on the choice of priors \cite{Gelman2008StatMed,Gelman2008StatAnnals,Gelman2020RegOther,Stan2020Priors,Gabry2023Priors}. The continuous adjustment variables are standardized to have zero mean and a standard deviation of 0.5. 
\begin{align*}
    \alpha &\sim \text{Student-t}_5(0, 10) \\
    \beta_2, \dots, \beta_6 &\sim \text{Student-t}_5(0, 2.5) \\
    \frac{1}{\sqrt{\phi}} &\sim \text{Half-Normal}(0,1),
\end{align*}
where Student-t$_5$ is a Student-t distribution with 5 degrees of freedom.

@@

@@

\begin{table}[!htpb]
    \caption[Estimation results from infection risk model]{Estimation results from infection risk model for the number of new respiratory cases.}
    \label{tab:estimation-results}
    \footnotesize
    \centering
    \begin{tabular}{lrrrrr}
    \toprule
    Parameter & Mean & Lower 95\%-CrI & Upper 95\%-CrI & $\hat{R}$ & ESS \\
    \midrule
    \input ../../results/epi-data/estimation-results
    \bottomrule
    \multicolumn{6}{p{13cm}}{\scriptsize
        ESS is the effective sample size, \ie the number of independent MCMC samples with estimation power equivalent to the total number of autocorrelated samples\cite{Stan2022}, and $\hat{R}$ is the Gelman-Rubin convergence diagnostic \cite{Gelman1992}. Low ESS or $\hat{R}$ or $\hat{R}>1.10$ indicate bad convergence of the model \cite{Gelman2013}.}
    \end{tabular}
\end{table}

@@

There can be unknown delay between when students were infected and when they tested positive. Therefore, we performed a sensitivity analysis where we lagged the molecular test results (number of positive saliva samples) by one test date (\ie from Tuesday to last Thursday and from Thursday to Tuesday) and one test week (\ie from Tuesday to last Tuesday and Thursday to last Thursday). The results are shown in \Cref{fig:mol-estimation-results-sensitivity}. The estimated effect of air cleaners remains indistinguishable from zero regardless of the considered lag.

\begin{figure}[!htb]
\centering
    \includegraphics[width=10cm]{../../results/mol-data/model-results.pdf}
    \caption[Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples]{Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples. Shown is the adjusted relative risk (posterior mean as dot and 50\%-, 80\%, and 95\%-CrIs as lines) for air cleaners by lag of the molecular test results.}
    \label{fig:mol-estimation-results-sensitivity}
\end{figure}

In the main paper, we compared particle concentrations between study conditions. \Cref{fig:env-descriptives-other-vars} compares additional environmental variables, \ie CO$_2$, air change rate, relative humidity, and temperature. The air change rate was estimated from indoor CO$_2$ levels using a transient mass balance model\cite{Batterman2017IJERPH}. There are no considerable differences between in any of the variables between study conditions. Furthermore, numerical estimation results for the reduction in particle concentrations with air cleaners are shown in \Cref{tab:palas-est-results}.

@@

>>>>

\title{\LARGE\singlespacing{\textbf{Appendix} \\ \medskip
Air cleaners and respiratory infections in schools: \\ A modeling study using epidemiological, environmental, and molecular data}}

@@

The clean air delivery rate (CADR) is typically used to quantify the cleaning efficiency of air cleaners.\cite{Nelson1993IndoorEnv} It is expressed in m$^3$/s and describes the volumetric flow rate of particle-free, clean air that an air cleaner delivers into the room. The CADR is equal to the product of the filtration efficiency and the volumetric flow rate passing through the device.

We followed the approach by K\"{u}pper et al. to determine the CADR.\cite{Kuepper20199AAQR} The approach is based on the release of fine aerosols (particle diameter $D$<1$\mu$m) in a closed, unventilated room by measuring the decay rate $k$ of the particle concentration $N(t)$ caused by the air cleaner, \ie
\begin{align}
    N(t) = N_0 \cdot e^{-k\cdot t}
\end{align}
Since the surfaces in the room (\eg walls and furniture) are also sinks for aerosol particles, measurements were carried out with ($k_\mathrm{cleaner}$) and without air cleaners($k_\mathrm{background}$). Based on that, we can compute the CADR as
\begin{align}
    \mathrm{CADR} = V \cdot k_\mathrm{cleaner} - V \cdot k_\mathrm{background},
\end{align}
where $V = 233 m^3$ is the volume of the classroom.

@@

Figure~\ref{fig:experiment} shows the particle number concentration with and without without the air cleaner. The exponential decrease is clearly visible as a linear curve in the logarithmic plot. Based on these experiments, we were able to determine the net CADR values of the air cleaner as 420\,m$^3$/h with an uncertainty of $\pm$10\% due to systematic differences in the setup and measurement errors of the instruments. Our estimated CADR is lower than the CADR reported by the manufacturer of 600m$^3$/h, most likely because the reported CADR was determined with comparably large particles (pollen in the super micrometer size range). Thus, we explain the lower filtration efficiency determined here with the fact that CADR is strongly dependent on the particle size, \ie large particles are easier to filter than fine aerosols.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{doc/paper/experiment-particle-concentration-decay.png}
    \caption[Measured concentration curves with and without air cleaner]{\textbf{Measured concentration curves with and without air cleaner.} The red dot shows the average concentration measured by the 11~distributed sensors of particles with $0.3\mu m < D < 1 \mu m$. The standard deviation is shown in gray as an error bar. The increase in concentration at the beginning of these experiments is due to the release of the NaCl particles during approximately 15-20\,min. The inhomogeneity of the concentration during the nebulization period is seen in larger error bars. Afterwards, the aerosol source is turned off, the concentration homogenizes quickly, and the concentration drop follows an exponential decay, which is used for the CADR calculation.}
    \label{fig:experiment}
\end{figure}

@@

\begin{table}[!htpb]
    \footnotesize
    \centering
    \caption[Overview of collected data]{\textbf{Overview of collected data}. Type of data collected, method/device, and frequency in the rooms of classes A and B.}
    \begin{tabular}{p{3.5cm}p{6cm} p{1cm} p{1cm} p{3cm}}
    \midrule
    Type of data & Method / device & Class A & Class B & Frequency \\
    \midrule
    \textbf{Molecular data} \\
    \midrule
    Bioaerosol sampling & Bioaerosol sampling devices: BioSpot-VIVAS (Aerosol Devices Inc., Ft. Collins, Colorado, United States) and Coriolis Micro Air (Bertin Instruments Montigny-le/Bretonneux, France) & x & (x) \newline only Coriolis & daily \\
    Saliva samples & Saliva samples & x & x & twice per week \\
    Swabs from air cleaners & Swab samples from Xiaomi HEPA filters (Xiaomi Mi Air Pro 70m2, Shenzhen, China) & x & x & after each intervention phase (both classes) and before the vacation (only class B) \\ 
    \midrule
    \textbf{Environmental data} \\
    \midrule
    CO$_2$, aerosol/particle concentrations, humidity, temperature & Air quality device: AQ Guard (Palas GmbH, Karlsruhe, Germany) & x & x & daily by minute \\
    Coughs & Sounds collected via microphones and coughs detected with a Machine Learning model\cite{Bertschinger2023CBMS} & x & x & daily by seconds \\
    \midrule
    \textbf{Epidemiological data} \\
    \midrule
    Absences & Survey & x & x & daily \\
    Student characteristics & Survey & x & x & at study start \\
    \bottomrule
    \end{tabular}
    \label{tab:data}
\end{table}

@@

Reports about absences were entered electronically into a REDcap database.\cite{Harris2009,Harris2019} \Cref{tab:epi-data-line-list} shows the line list data for respiratory cases based on epidemiological data. For each class and case, it shows the date when the student was absent, s/he reported symptoms, and returned to school. We also collected data on laboratory tests but none of the students reported a test result during the study. A case of respiratory infection was defined as an absence where the student reported a sickness with at least one of the following symptoms: fever, coughing, tiredness, loss of test or smell, sore throat, headache, aches and pains, diarrhea, difficulty breathing or shortness of breath, stomach. Students were asked to report the first date they experienced symptoms, so that a student absent on Monday could report Saturday or Sunday as the day when symptoms began to be felt. We will refer respiratory cases by their dates of symptom onset, which usually corresponded to the absence date, unless the student attended the school while experiencing symptoms. Informed consent for one student could not be obtained and the single absence of this student was considered unrelated to a respiratory infection.

@@

The prior distributions for the virus-specific incubation periods are based on estimates published in the literature.\cite{McAloon2020,Lessler2009LancetID} The distributions are shown in \Cref{fig:prior-pin}. Since we could not obtain prior estimates for the incubation period of metapneumovirus (MPV) from the literature, we instead formed a distribution by using the equally weighted average of the parameters from the other distributions.

@@

We use weakly informative priors for all modeling parameters following recommendations on the choice of priors.\cite{Gelman2008StatMed,Gelman2008StatAnnals,Gelman2020RegOther,Stan2020Priors,Gabry2023Priors} The continuous adjustment variables are standardized to have zero mean and a standard deviation of 0.5. 
\begin{align*}
    \alpha &\sim \text{Student-t}_5(0, 10) \\
    \beta_2, \dots, \beta_6 &\sim \text{Student-t}_5(0, 2.5) \\
    \frac{1}{\sqrt{\phi}} &\sim \text{Half-Normal}(0,1),
\end{align*}
where Student-t$_5$ is a Student-t distribution with 5 degrees of freedom.

@@

In the main paper, we compared particle concentrations between study conditions. \Cref{fig:env-descriptives-other-vars} compares additional environmental variables, \ie CO$_2$, air change rate, relative humidity, and temperature. The air change rate was estimated from indoor CO$_2$ levels using a transient mass balance model.\cite{Batterman2017IJERPH} There are no considerable differences between in any of the variables between study conditions. Furthermore, numerical estimation results for the reduction in particle concentrations with air cleaners are shown in \Cref{tab:palas-est-results}.

@@

\begin{table}[!htpb]
    \caption[Estimation results from infection risk model]{Estimation results from infection risk model for the number of new respiratory cases.}
    \label{tab:estimation-results}
    \footnotesize
    \centering
    \begin{tabular}{lrrrrr}
    \toprule
    Parameter & Mean & Lower 95\%-CrI & Upper 95\%-CrI & $\hat{R}$ & ESS \\
    \midrule
    \input ../../results/epi-data/estimation-results
    \bottomrule
    \multicolumn{6}{p{13cm}}{\scriptsize
        ESS is the effective sample size, \ie the number of independent MCMC samples with estimation power equivalent to the total number of autocorrelated samples,\cite{Stan2022} and $\hat{R}$ is the Gelman-Rubin convergence diagnostic.\cite{Gelman1992} Low ESS or $\hat{R}$ or $\hat{R}>1.10$ indicate bad convergence of the model.\cite{Gelman2013}}
    \end{tabular}
\end{table}

@@

@@

There can be unknown delay between when students were infected and when they tested positive. Therefore, we performed a sensitivity analysis where we lagged the molecular test results (number of positive saliva samples) by one test date (\ie from Tuesday to last Thursday and from Thursday to Tuesday) and one test week (\ie from Tuesday to last Tuesday and Thursday to last Thursday). The results are shown in \Cref{fig:mol-estimation-results-sensitivity}. The estimated effect of air cleaners remains indistinguishable from zero regardless of the considered lag.  
\begin{figure}[!htb]
\centering
    \includegraphics[width=10cm]{../../results/mol-data/model-results.pdf}
    \caption[Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples]{Sensitivity analysis for the estimated effect of air cleaners on the positivity rate of saliva samples. Shown is the adjusted relative risk (posterior mean as dot and 50\%-, 80\%, and 95\%-CrIs as lines) for air cleaners by lag of the molecular test results.}
    \label{fig:mol-estimation-results-sensitivity}
\end{figure}
